---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


```{r setup, include=FALSE, message= FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Loading and preprocessing the data


```{r read in data}
df <- read.csv(paste0('/Users/garymu/Dropbox (Personal)/Coursera/DS/',
               'reproducible_research/assignment1/activity.csv'))
```

## What is mean total number of steps taken per day?

```{r steps_per_day}
summary_steps <- df %>% 
    filter(!is.na(steps)) %>%
    group_by(date) %>%
    summarise(total_steps = sum(steps, na.rm = T),
              mean_steps = mean(steps, na.rm = T),
              median_steps = median(steps, na.rm = T))

p1 <- ggplot(summary_steps) + 
        geom_histogram(aes(x = total_steps)) +
        ggtitle('Total Steps Per Day')

#histogram of total steps taken per day
p1

#summary of steps taken per day (total, mean, median)
summary_steps

#mean, median total number of steps taken per day
print(paste0('mean total number of steps taken per day is: ', mean(summary_steps$mean_steps, na.rm = T)))
print(paste0('median total number of steps taken per day is: ',median(summary_steps$mean_steps, na.rm = T)))
```

mean total number of steps taken per day is 37.3826 steps, and median is 37.37847 steps.

## What is the average daily activity pattern?

```{r steps_over_5_min_interval}

interval <- df %>% 
                group_by(interval) %>%
                summarise(mean_steps = mean(steps, na.rm = T))

p2 <- ggplot(interval, aes(interval, mean_steps)) + 
        geom_line() +
        ggtitle('Mean Steps Per Interval Across Dates')

#histogram of mean steps taken per interval Across Dates
p2

#which interval has the highest mean steps
interval$interval[which.max(interval$mean_steps)]
```
The interval that has the highest mean steps is interval 835.

## Imputing missing values

```{r Imputing_missing_values, message= FALSE}
#total dimension of steps data
dim(df)

#total number of rows with missing steps
sum(is.na(df$steps))

#impute the NA rows with the mean value of the day

df2 <- df

for(i in 1:nrow(df2)){
    if(is.na(df2$steps[i])){
        if(df2$date[i] %in% as.character(summary_steps$date)){
             df2$steps[i] = summary_steps$mean_steps[summary_steps$date== df2$date[i]]
        }else{
            df2$steps[i] = 0
        }
    }
}


summary_steps2 <- 
    df2 %>% 
        group_by(date) %>% 
        summarise(total_steps = sum(steps),
                  mean_steps = mean(steps))

p3 <- ggplot(data = summary_steps2) + 
              geom_histogram(aes(x = total_steps))+
              ggtitle('Total Steps Per Day (imputed data)')


p3

#original steps mean
mean(df$steps, na.rm = T)

#steps mean with na imputed with the mean of the day
mean(df2$steps)
```
There are 2304 observations with NA values.


The mean increased after we imputed values as expected: 32.48 vs 37.38.

## Are there differences in activity patterns between weekdays and weekends?

```{r Mean_Steps_Per_Interval_by_Weekend_Weekdays }
#convert the dates to weekday
df2 <- df2 %>% 
            mutate(weekday = weekdays(as.Date(date))) %>%
            mutate(weekday2 = ifelse(weekday %in% c('Sunday', 'Saturday'), 'weekend', 'weekday'))

df3 <- df2 %>%
    group_by(weekday2, interval) %>%
    summarise(mean_steps = mean(steps))

ggplot(df3)+ 
    geom_line(aes(x = interval, y = mean_steps)) + 
    facet_wrap(~weekday2, ncol = 1) +
    ggtitle('Mean Steps Per Interval by Weekend/Weekdays')

```

